Memory for television episodes preserves event content while introducing new across-event similarities

### 2020-02-04 paper/main.tex
some updates to results section

\cite{ChenEtal17} collected hand-annotated information about each of 1000 (manually identified) scenes spanning the roughly 50 minute video used in their experiment.  This information included: a brief narrative description of what was happening; the location where the scene took place; the names of any characters on the screen; and other similar details (for a full list of annotated features, see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and," ``or," ``but," etc.) across all features and scenes as the ``vocabulary" for the topic model.  We then concatenated the sets of words across all features contained in overlapping, sliding windows of (up to) 50 scenes, and treated each window as a single ``document" for the purpose of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 32 unique topics (with non-zero weights) were sufficient to describe the time-varying content of the video (see \textit{Methods}; Figs.~\ref{fig:schematic}, \topics).  Note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06} in that we sought to characterize how the thematic content of the episode evolved over time.  However, whereas Dynamic Topic Models are designed to characterize how the properties of \textit{collections} of documents change over time, our sliding window approach allows us to examine the topic dynamics within a single document (or video).  Specifically, our approach yielded (via the topic proportions matrix) a single \textit{topic vector} for each sliding window of annotations transformed by the topic model.  We then stretched the resulting windows-by-topics matrix to match the time series of the 1976 fMRI volumes collected as participants viewed the episode.

@@

Statistical models of memory studies often treat recall success as binary (i.e., an item either was or was not recalled), or occasionally categorical \citep[e.g., to distinguish familiarity from recollection][]{YoneEtal02}.  Such approaches are tenable in classical list-learning or recognition memory paradigms, as the presented stimuli tend to be very simple (e.g., a sequence of individual words or items).  However, the feature-rich content of a naturalistic experiences may later be described with many, highly variable levels of success.  Our framework produces a content-based model of individual stimulus and recall events, allowing for direct quantitative comparison between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel, \textit{continuous} metrics for quantifying naturalistic memory representations:~\textit{precision} and \textit{distinctiveness}.  We define precision as the ``completeness" of recall-- i.e., how fully the presented content was recapitulated in memory.  Under our framework, we quantify this for a given recall event as the correlation between the topic proportions of the recall event and the maximally correlated video event (Fig.~\ref{fig:precision-distinctiveness}).  A second novel metric we introduce here is \textit{distinctiveness}, which we define as the ``specificity" of recall-- i.e., how unique the description of a given section of content was, compared to descriptions for other sections of content.  We quantify this for each recall event as 1 minus the average correlation between the given recall event and all other recall events not matched to the same video event.  In addition to individual events, one may also use these metrics to describe each participant's overall performance.  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that, across participants, a higher precision score is correlated to both hand-annotated memory performance (Pearson's $r(15) = 0.56, p = 0.021$) and the number of recall events estimated by our model (Pearson's $r(15) = 0.85, p < 0.001$).  We also hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  We find that this distinctiveness score is related to our model's estimated number of recalled events (Pearson's $r(15) = 0.53, p = 0.028$).  While we do not find distinctiveness to be related to hand-annotated memory performance (Pearson's $r(15) = 0.28, p = 0.275$), this is not entirely surprising given how the hand-annotated memory scores were computed (see \textit{Methods}).

@@

Further intuition for the behaviors captured by these two metrics may be gained by directly examining the content of the video and recalls our framework models.  In Figure \ref{fig:precision-detail}, we contrast two participants' recall descriptions of the same video event: one with a high precision score, the other with a low precision score.  From the HMM-identified event boundaries, we recovered the set of annotations describing the content of an example video event (panel B), and divided them into different color-coded sections for each action or feature described.  We then similarly recovered the set of sentences comprising the corresponding recall event for each of the two example participants.  Because the recall sliding windows overlap heavily, and each recall event spans multiple recall timepoints (i.e., windows), we have stripped any sentences from the beginning and end that describe earlier or later video events.  In other words, the sentences displayed in Fig.~\ref{fig:precision-detail}C are a subset of the full recall event text, spanning sentences between the first and last descriptions of content from the example video event.  We then colored all descriptions of the video event's content by the

@@

\subsection*{Experimental design and data collection}
Data were collected by \cite{ChenEtal17}.  In brief, participants ($n=17$) viewed the first 48 minutes of ``A Study in Pink'', the first episode of the BBC television series \textit{Sherlock}, while fMRI volumes were collected (TR = 1500~ms).  The stimulus was divided into a 23~min (946~TR) and a 25~min (1030~TR) segment to mitigate technical issues related to the scanner.  After finishing the clip, participants were instructed to \citep[quoting from][]{ChenEtal17} ``describe what they recalled of the [episode] in as much detail as they could, to try to recount events in the original order they were viewed in, and to speak for at least 10 minutes if possible but that longer was better. They were told that completeness and detail were more important than temporal order, and that if at any point they realized they had missed something, to return to it. Participants were then allowed to speak for as long as they wished, and verbally indicated when they were finished (e.g., `I’m done').''  For additional details about the experimental procedure and scanning parameters, see \cite{ChenEtal17}.  The experimental protocol was approved by Princeton University's Institutional Review Board.

@@

@@

\subsection*{Statistics}
All statistical tests we performed were two-sided.

>>>>

\cite{ChenEtal17} collected hand-annotated information about each of 1000 (manually identified) time segments spanning the roughly 50 minute video used in their experiment.  This information included: a brief narrative description of what was happening, the location where the scene took place, the names of any characters on the screen, and other similar details (for a full list of annotated features, see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and," ``or," ``but," etc.) across all features and scenes as the ``vocabulary" for the topic model.  We then concatenated the sets of words across all features contained in overlapping, sliding windows of (up to) 50 scenes, and treated each window as a single ``document" for the purpose of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 32 unique topics (with non-zero weights) were sufficient to describe the time-varying content of the video (see \textit{Methods}; Figs.~\ref{fig:schematic}, \topics).  Note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06} in that we sought to characterize how the thematic content of the episode evolved over time.  However, whereas Dynamic Topic Models are designed to characterize how the properties of \textit{collections} of documents change over time, our sliding window approach allows us to examine the topic dynamics within a single document (or video).  Specifically, our approach yielded (via the topic proportions matrix) a single \textit{topic vector} for each sliding window of annotations transformed by the topic model.  We then stretched the resulting windows-by-topics matrix to match the time series of the 1976 fMRI volumes collected as participants viewed the episode.

@@

Statistical models of memory studies often treat recall success as binary (i.e., an item either was or was not recalled), or occasionally categorical \citep[e.g., to distinguish familiarity from recollection;][]{YoneEtal02}.  Such approaches are tenable in classical list-learning or recognition memory paradigms, as the presented stimuli tend to be very simple (e.g., a sequence of individual words or items).  However, the feature-rich content of a naturalistic experiences may later be described with many, highly variable levels of success.  Our framework produces a content-based model of individual stimulus and recall events, allowing for direct quantitative comparison between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel, \textit{continuous} metrics for quantifying naturalistic memory representations:~\textit{precision} and \textit{distinctiveness}.  We define precision as the ``completeness" of recall, or how fully the presented content was recapitulated in memory.  Under our framework, we quantify this for a given recall event as the correlation between the topic proportions of the recall event and the maximally correlated video event (Fig.~\ref{fig:precision-distinctiveness}).  A second novel metric we introduce here is \textit{distinctiveness}, which we define as the ``specificity" of recall, or how unique the description of a given section of content was, compared to descriptions for other sections of content.  We quantify this for each recall event as 1 minus the average correlation between the given recall event and all other recall events not matched to the same video event.  In addition to individual events, one may also use these metrics to describe each participant's overall performance (i.e., by averaging across a participant's event-wise precision or distinctiveness scores).  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that, across participants, a higher precision score is correlated to both hand-annotated memory performance (Pearson's $r(15) = 0.56, p = 0.021$) and the number of recall events estimated by our model (Pearson's $r(15) = 0.85, p < 0.001$).  We also hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  We find that this distinctiveness score is related to our model's estimated number of recalled events (Pearson's $r(15) = 0.53, p = 0.028$), and while we do not find distinctiveness to be related to hand-annotated memory performance (Pearson's $r(15) = 0.28, p = 0.275$), this is not entirely surprising given how the hand-annotated memory scores were computed (see \textit{Methods}).

@@

Further intuition for the behaviors captured by these two metrics may be gained by directly examining the content of the video and recalls our framework models.  In Figure \ref{fig:precision-detail}, we contrast recalls for the same video event from two participants: one with a high precision score (P17), the other with a low precision score (P6).  From the HMM-identified event boundaries, we recovered the set of annotations describing the content of an example video event (event 22; Fig.~\ref{fig:precision-detail}B), and divided them into different color-coded sections for each action or feature described.  We then similarly recovered the set of sentences comprising the corresponding recall event for each of the two example participants.  Because the recall sliding windows overlap heavily, and each recall event spans multiple recall timepoints (i.e., windows), we have stripped any sentences from the beginning and end that describe earlier or later video events for the sake of readability.  In other words, Fig.~\ref{fig:precision-detail}C shows a subset of the full recall event text, comprising sentences between the first and last descriptions of content from the example video event.  We then colored all words describing actions and features coded in panel B by their corresponding color.  Visual comparison of the transcripts reveals that the most precise participant's recall captures a greater amount of the video event's content, and with far more detail.

Figure \ref{fig:distinctiveness-detail} similarly contrasts two participants' recall for an example video event (event 19) to illustrate the tangible differences between high and low distinctiveness scores.  Here, we have extracted the full text from the most

@@

\subsection*{Experimental design and data collection}
Data were collected by \cite{ChenEtal17}.  In brief, participants ($n=22$) viewed the first 48 minutes of ``A Study in Pink'', the first episode of the BBC television series \textit{Sherlock}, while fMRI volumes were collected (TR = 1500~ms).  Participants were pre-screened to ensure they had never seen any episode of the show before.  The stimulus was divided into a 23~min (946~TR) and a 25~min (1030~TR) segment to mitigate technical issues related to the scanner.  After finishing the clip, participants were instructed to \citep[quoting from][]{ChenEtal17} ``describe what they recalled of the [episode] in as much detail as they could, to try to recount events in the original order they were viewed in, and to speak for at least 10 minutes if possible but that longer was better. They were told that completeness and detail were more important than temporal order, and that if at any point they realized they had missed something, to return to it. Participants were then allowed to speak for as long as they wished, and verbally indicated when they were finished (e.g., `I’m done').''  Five participants were dropped from the original dataset due to excessive head motion (2 participants), insufficient recall length (2 participants), or falling asleep during stimulus viewing (1 participant), resulting in a final sample size of $n=17$.  For additional details about the experimental procedure and scanning parameters, see \cite{ChenEtal17}.  The experimental protocol was approved by Princeton University's Institutional Review Board.

@@

The video stimulus was divided into 1,000 fine-grained ``scenes" and annotated by an independent coder.  For each of these 1,000 scenes, the following information was recorded: a brief narrative description of what was happening, the location where the scene took place, whether that location was indoors or outdoors, the names of all characters on-screen, the name(s) of the character(s) in focus in the shot, the name(s) of the character(s) currently speaking, the camera angle of the shot, a transcription of any text appearing on-screen, and whether or not there was music present in the background.  Each scene was also tagged with its onset and offset time, in both seconds and TRs.

The video was also divided by an independent coder into 50 more broad ``scenes" ``following major shifts in the narrative (e.g., location, topic, and/or time)" \citep{ChenEtal17}.  The hand-annotated memory scores for each participant we reference in our present study were generated by considering a scene to have been recalled (in a binary fashion) ``if the participant described any part of the scene."

@@

\subsection*{Statistics}
All statistical tests performed in the behavioral analyses were two-sided.  All statistical tests performed in the neural data analyses were two-sided, except for the permutation-based thresholding, which was one-sided.  In this case, we were specifically interested in identifying voxels whose activation time series reflected the temporal structure of the video and recall trajectories to a \textit{greater} extent than that of the phase-shifted trajectories.
% ------ NOTE: ------
% from the Nature Human Behavior submission requirements: "Specify whether tests were one- or two-tailed, and justify the use of one-tailed tests."
% - Paxton
